### 2016-03-22 paper.tex
Adopt 'culture-narration' as new genre title

\newcommand{\projecttitle}{Definition and directions of narrative-rich games}

@@

% Abstract: 150 words or less, extra 0.5in left and right margin.
% Note that all text is justified.
%
\begin{adjustwidth}{0.5in}{0.5in}
  \textbf{Abstract:} \textit{Player-narration games} are a promising
  genre for transformative games. This genre can be defined formal and
  dramatic characteristics, and we identify tabletop games that belong
  to the genre.  Adding technological components to the game affords
  new interaction modes as well as convenient digital distribution. We
  describe our team's pilot project at creating a technology-enhanced
  player-narration game to improve cultural empathy among players aged
  10--14.  This pilot project highlights several complications with
  this genre, including thematic, technical, and production
  challenges.  However, the pilot project justifies continued
  experimentation, development, and formal assessment within the
  genre.
\end{adjustwidth}

@@

\section{Defining the genre of player-narration games}

We propose that \totan{} is one of very few games comprising a genre
of \textit{player-narration games}, and we define this genre as having
the following characteristics.

@@

For contrast, consider \textit{Above and Below}~\citep{Laukat2015}, a
worker-management game about building a city both above and below the
surface.  \textit{Above and Below} is primarily about building an
economic engine that generates victory points faster than your
opponents within a fixed number of rounds. Exploring the underground
is one source of points, and this is done by determining a random
entry within its encounter book.  A player reads this entry aloud,
after which the player is presented with a decision of how to
react. However, the reaction titles are a thin wrapper around a
risk-and-reward balance: each has a number which is the number of
successes required on dice rolls, so players choose the highest number
they believe they can roll. Higher numbers always equate to better results,
and this roll is usually the end of the encounter: a successful player
is given an in-game reward with no other story or narrative.
Hence, although \textit{Above and Below} includes an encounter book
that players read to each other, it does not satisfy the characteristics
to be in the player-narration game genre.

@@

Investigating the potential for player-narration games to teach
cultural empathy brought our team into a partnership with
\textit{a local library branch hidden for blind review}. This branch
serves primarily patrons from low socio-economic status, although
its special technology education programs also draw homeschoolers from
around the region.
The youth patrons of the library tend to be 4rd--6th grade 
``latchkey kids'' who 
see the library as a place to visit with friends, to participate in programs,
and to access technology.

Conversations with program coordinators about player-narration games
revealed that despite other differences, the youth patrons shared
a fascination with monsters, horror, and the macabre. 
This is in part because of cultural taboo: the youth will brag about
being able to see a PG-13-rated movie or play an M-rated game,
for example, and several kids worked together to produce an original
short movie based on battles between movie monsters.

The fascination exhibited by the youth patrons reflects various theses
of Cohen's monster theory~\citep{Cohen1996}. 
We see in their behavior that ``the same creatures who terrify and
interdict can evoke potent escapist fantasies''~(p.16).
Recognizing the youth patrons' attraction to this theme,
Cohen's theory establishing monsters as reflections of culture,
and the narrative potential of monsters,
our team decided to engage in a design experiment to validate our
hypotheses about player-narration games,
in collaboration with \textit{the library branch}.

>>>>

\newcommand{\projecttitle}{Definition and directions of cultural-narrative games}

@@

% Abstract: 150 words or less, extra 0.5in left and right margin.
% Note that all text is justified.
%
\begin{adjustwidth}{0.5in}{0.5in}
  \textbf{Abstract:} \textit{Culture-narration games} are a promising
  genre for transformative games. This genre can be defined formal and
  dramatic characteristics, and we identify tabletop games that belong
  to the genre.  Adding technological components to the game affords
  new interaction modes as well as convenient digital distribution. We
  describe our team's pilot project at creating a technology-enhanced
  culture-narration game to improve cultural empathy among players aged
  10--14.  This pilot project highlights several complications with
  this genre, including thematic, technical, and production
  challenges.  However, the pilot project justifies continued
  experimentation, development, and formal assessment within the
  genre.
\end{adjustwidth}

@@

\section{Defining the genre of culture-narration games}

We propose that \totan{} is one of very few games comprising a genre
of \textit{culture-narration games}, and we define this genre as having
the following characteristics.

@@

For contrast, consider \textit{Above and Below}~\citep{Laukat2015}, a
worker-management game about building a city both above and below the
surface.  \textit{Above and Below} is primarily about building an
economic engine that generates victory points faster than your
opponents within a fixed number of rounds. Exploring the underground
is one source of points, and this is done by determining a random
entry within its encounter book.  A player reads this entry aloud,
after which the player is presented with a decision of how to
react. However, the reaction titles are a thin wrapper around a
risk-and-reward balance: each has a number which is the number of
successes required on dice rolls, so players choose the highest number
they believe they can roll. Higher numbers always equate to better results,
and this roll is usually the end of the encounter: a successful player
is given an in-game reward with no other story or narrative.
Hence, although \textit{Above and Below} includes an encounter book
that players read to each other, it does not satisfy the characteristics
to be in the culture-narration game genre.

@@

Investigating the potential for culture-narration games to teach
cultural empathy brought our team into a partnership with
\textit{a local library branch hidden for blind review}. This branch
serves primarily patrons from low socio-economic status, although
its special technology education programs also draw homeschoolers from
around the region.
The youth patrons of the library tend to be 4rd--6th grade 
``latchkey kids'' who 
see the library as a place to visit with friends, to participate in programs,
and to access technology.

Conversations with program coordinators about culture-narration games
revealed that despite other differences, the youth patrons shared
a fascination with monsters, horror, and the macabre. 
This is in part because of cultural taboo: the youth will brag about
being able to see a PG-13-rated movie or play an M-rated game,
for example, and several kids worked together to produce an original
short movie based on battles between movie monsters.

The fascination exhibited by the youth patrons reflects various theses
of Cohen's monster theory~\citep{Cohen1996}. 
We see in their behavior that ``the same creatures who terrify and
interdict can evoke potent escapist fantasies''~(p.16).
Recognizing the youth patrons' attraction to this theme,
Cohen's theory establishing monsters as reflections of culture,
and the narrative potential of monsters,
our team decided to engage in a design experiment to validate our
hypotheses about culture-narration games,
in collaboration with \textit{the library branch}.
